### 2020-03-26 Note.Rmd
add note

# Note for the paper

ERGM toturial: 
https://www.r-bloggers.com/ergm-tutorial/

Bergm toturial:
http://acaimo.github.io/Bergm/tutorials/Dec_18.html

# Explore different combination of variables for Bergm
1. there is difference in statistics (mean, sd) between weighed and unweighed ties
2. difference in statistics when simplifying variables to agate and golden beads, plus a few types of prestige goods such as gold foil and fish-like shaped necklace. The mean looks fine
3. did not work well when there is too many ties in the network
4. difference in the results of using BergmM and Bergm, BergmM is for missing data
5. add the spacial information?